---
title: "GRanges"
author: "tkik"
date: "2023-06-07"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


## Granges and GRangesList

`GRanges` objects store information about Genomic ranges with chromosome, position and additional information. Offers a wide range of functionality.
`GRangesList` is a "grouped" version of `GRnages`, it is similar to a normal list, but with additional functionality. 
Both of them are able to handle strand information as well.

```{r}
library(GenomicRanges)

gr1 <- GRanges(
    seqnames = "chr2",
    ranges = IRanges(103, 106),
    strand = "+",
    score = 5L, GC = 0.45)
gr2 <- GRanges(
    seqnames = c("chr1", "chr3"),
    ranges = IRanges(c(107, 113), width = 3),
    strand = c("+", "-"),
    score = 3:4, GC = c(0.3, 0.5))
gr3 <- GRanges("chr2:102-107")

gr4 <- makeGRangesFromDataFrame(data.frame(chr=c("chr1", "chr2"), 
                                           start=c(104, 104), end=c(108, 105),
                                           name=c("gene1", "gene2")), keep.extra.columns = T, ignore.strand = T)
grl <- GRangesList("txA" = gr1, "txB" = gr2)
grl

#seqinfo(gr1) <- Seqinfo(genome="hg38")
#can have names for each row


```


### Accessors

```{r}
seqnames(gr1)
seqlevels(gr1)
ranges(gr1)
start(gr1)
end(gr1)
strand(gr1)
width(gr1)
length(gr1)
mcols(gr4)
mcols(gr4)$name

```

### Subsetting


```{r}
#data.frame-like

gr1[1,]
gr4[1, "name"]


# subset function
subset(gr1, strand == "+")

```

### Splitting and combining

```{r}

sp <- split(gr2, 1:2)
split(gr4, ~ name)

c(sp[[1]], sp[[2]])

stack(sp, index.var="name")


```

### Aggregating

```{r}
aggregate(gr2, score ~ strand, mean)

```

### Interval operations

```{r}
flank(gr1, 10, both=T)
shift(gr2, 5)
resize(gr1, 20)
c(gr1, gr3)
 reduce(c(gr1, gr3)) # combines the overlapping ranges

 reduce(c(gr1, gr3), ignore.strand=T)
```

### Comapring 2 GRanges

```{r}
union(gr2, gr4)

intersect(gr2, gr4)
setdiff(gr2, gr4)

hits <- findOverlaps(gr2, gr4, ignore.strand=TRUE)
countOverlaps(gr2, gr4, ignore.strand=TRUE)
subsetByOverlaps(gr2, gr4)


```

# Exercises 

1. Create a GRanges object using the information in the table below:
chr	start	end	strand	score
chr1	10000	10300	+	10
chr1	11100	11500	-	20
chr2	20000	20030	+	15

List the start, end, strand and chomrosomes of the generated object. 

```{r}
gr1 <- GRanges(c("chr1:10000-10300:+", "chr1:11100-11500:-", "chr2:20000-20030:+"))
mcols(gr1)$score <- c("10", "20", "15")

start(gr1)
end(gr1)
strand(gr1)
seqnames(gr1)
```


Use the start(), end(), strand(),seqnames() and width() functions on the GRanges object you created.
1. Subset only the ones on the + strand. 
2. subset the ones on the chr1

```{r}
gr1[strand(gr1)=="+",]
gr1[seqnames(gr1)=="chr1"]
```


Download the coordinates of the promoters of the hg38 genome. 

```{r}
library(annotatr)
annots = c('hg38_genes_promoters')
annots_gr = build_annotations(genome = 'hg38', annotations = annots)


```

Open the provided CUT&TAG data for SMAD3

```{r}
library(rtracklayer)

smad3 <- import.bed("data/GSM7446496_SMAD3_U3008_01_noIgG.bed", genome="hg38")

```

Find the peaks that are overlapping with a promoter. See if there are more overlaps than expected. In order to do this, generate a random regions. 


```{r}

ol <- findOverlaps(smad3, annots_gr)


smad3_random <- randomize_regions(
  smad3,
  allow.overlaps = FALSE,
  per.chromosome = TRUE,
  quiet = FALSE
)

cl <- countOverlaps(smad3, annots_gr)
cl_random <- countOverlaps(smad3_random, annots_gr)

cont <- matrix(c(sum(cl!=0), sum(cl_random!=0), length(smad3)-sum(cl!=0), length(smad3)-sum(cl_random!=0)), nrow=2)

chisq.test(cont)

```

